---
title: "Statistical Inference Course Project (Part 2)"
author: "BELMOUIDI Mohamed"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: false
    df_print: kable
    number_sections: false
    fig_caption: yes
    highlight: tango
    dev: pdf
---

## Course Project

**Basic Inferential Data Analysis**

### Instructions

* Load the ToothGrowth data and perform some basic exploratory data analyses
* Provide a basic summary of the data.
* Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there are other approaches worth considering.)
* State your conclusions and the assumptions needed for your conclusions.

---

### Exploratory Data Analysis

First, we load the required packages and the dataset:

```{r, echo=FALSE}
# Load required packages
library(dplyr, warn.conflicts = FALSE)
library(ggplot2)
library(ggthemes)

# Load data and convert to tbl format
ToothGrowth <- tbl_df(ToothGrowth)
```

We examine the structure and summarize the dataset:

```{r, echo=FALSE}
# Structure of the dataframe
str(ToothGrowth)

# Summary
summary(ToothGrowth)
```

We observe the unique values of the `dose` variable:

```{r, echo=FALSE}
# Unique values in the dose vector
unique(ToothGrowth$dose)
```

The `dose` variable contains three unique values (0.5, 1, 2). We convert it to a factor:

```{r, echo=FALSE}
# Convert dose to a factor
ToothGrowth <- ToothGrowth %>% mutate(dose = as.factor(dose))
```

#### Plots

```{r, echo=FALSE, fig.height=3}
ToothGrowth %>%
  ggplot(aes(x = dose, y = len, fill = supp)) +
  geom_boxplot() +
  facet_grid(. ~ supp) +
  scale_fill_brewer(palette = "Set1") +
  theme_bw() +
  ggtitle("Teeth Length vs Dose Level \nby Supplement Type") +
  labs(x = "Dose (mg)", y = "Teeth Length") +
  guides(fill = guide_legend(title = "Supplement Type"))
```

```{r, echo=FALSE, fig.height=3}
ToothGrowth %>%
  ggplot(aes(x = supp, y = len)) +
  geom_boxplot(aes(fill = supp)) +
  facet_wrap(~ dose) +
  scale_fill_brewer(palette = "Set1") +
  theme_bw() +
  ggtitle("Teeth Length vs Supplement Type \nby Dose Level") +
  labs(x = "Supplement Type", y = "Teeth Length") +
  guides(fill = guide_legend(title = "Supplement Type"))
```

---

### Hypothesis Tests

We conduct statistical tests to compare teeth growth by supplement type and dose levels. 

#### Test by Dose Levels

* ##### Test A: Dose = 0.5 and Dose = 1

```{r, echo=FALSE}
len_a <- ToothGrowth %>% filter(dose %in% c(0.5, 1)) %>% pull(len)
dose_a <- ToothGrowth %>% filter(dose %in% c(0.5, 1)) %>% pull(dose)
Test.a <- t.test(len_a ~ dose_a)
Test.a
```

* ##### Test B: Dose = 0.5 and Dose = 2

```{r, echo=FALSE}
len_b <- ToothGrowth %>% filter(dose %in% c(0.5, 2)) %>% pull(len)
dose_b <- ToothGrowth %>% filter(dose %in% c(0.5, 2)) %>% pull(dose)
Test.b <- t.test(len_b ~ dose_b)
Test.b
```

* ##### Test C: Dose = 1 and Dose = 2

```{r, echo=FALSE}
len_c <- ToothGrowth %>% filter(dose %in% c(1, 2)) %>% pull(len)
dose_c <- ToothGrowth %>% filter(dose %in% c(1, 2)) %>% pull(dose)
Test.c <- t.test(len_c ~ dose_c)
Test.c
```

#### Test by Supplement Type

```{r, echo=FALSE}
len <- ToothGrowth %>% pull(len)
supp <- ToothGrowth %>% pull(supp)
Test.supp <- t.test(len ~ supp)
Test.supp
```

---

### Conclusions

**Key Findings:**

1. There is a statistically significant relationship between dose levels and teeth length (p-values < 0.05 in all dose comparisons).
2. No statistically significant difference exists between supplement types (p-value = 0.06).

**Assumptions for Validity:**

- Independence of samples: Random sampling or assignment.
- Normality of population distributions.

These assumptions are critical to ensure the validity of t-test results.
